---
title: "Late Round Prospect Guide — Evaluation 2022–2025"
author: "Nick Gurol"
output:
  html_document:
    toc: true
    toc_float: true
    self_contained: true
knit: (function(input, ...) rmarkdown::render(input, output_dir = dirname(input), ...))
---

```{css mobile-styles, echo=FALSE}
img { max-width: 100%; height: auto; }
.main-container { max-width: 1000px; }
table { display: block; overflow-x: auto; }
@media (max-width: 768px) {
  .main-container { max-width: 100% !important; padding: 8px !important; }
  table { font-size: 11px !important; }
  h1 { font-size: 1.5em; } h2 { font-size: 1.25em; } h3 { font-size: 1.1em; }
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 9, fig.height = 5, dpi = 120)
suppressPackageStartupMessages({
  library(tidyverse)
  library(arrow)
  library(knitr)
  library(jsonlite)
})

# Resolve repo root robustly. Three strategies, in order:
#   1) rstudioapi (only works inside RStudio)
#   2) knitr current input path (set when knitting)
#   3) getwd() fallback — assumes we're somewhere inside the repo
find_repo_root <- function() {
  candidates <- character(0)

  rs_path <- tryCatch(
    rstudioapi::getActiveDocumentContext()$path,
    error = function(e) NULL
  )
  if (!is.null(rs_path) && nzchar(rs_path)) {
    candidates <- c(candidates, normalizePath(file.path(dirname(rs_path), "..", ".."),
                                              mustWork = FALSE))
  }

  knit_input <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) NULL)
  if (!is.null(knit_input) && nzchar(knit_input)) {
    candidates <- c(candidates, normalizePath(file.path(dirname(knit_input), "..", ".."),
                                              mustWork = FALSE))
  }

  # If knitting, working dir is the Rmd dir → ../.. is repo root.
  candidates <- c(candidates, normalizePath(file.path(getwd(), "..", ".."), mustWork = FALSE))
  # If working dir is already repo root.
  candidates <- c(candidates, normalizePath(getwd(), mustWork = FALSE))

  marker <- "analysis/late_round_eval/data/eval_df.parquet"
  for (cand in candidates) {
    if (file.exists(file.path(cand, marker))) return(cand)
  }
  # Last resort: return first candidate even if marker missing
  candidates[1]
}

repo_root <- find_repo_root()
# Make repo_root visible to later chunks (knitr resets cwd between chunks)
knitr::opts_knit$set(root.dir = repo_root)
setwd(repo_root)

# Absolute paths so later chunks don't depend on cwd
abs_path <- function(rel) file.path(repo_root, rel)

source(abs_path("analysis/late_round_eval/data_pipeline.R"), local = TRUE)
source(abs_path("analysis/late_round_eval/models.R"), local = TRUE)
source(abs_path("analysis/late_round_eval/charts.R"), local = TRUE)

# Re-anchor chart output dir to absolute path (charts.R sets a relative CHART_DIR)
CHART_DIR <- abs_path("analysis/late_round_eval/charts")
dir.create(CHART_DIR, recursive = TRUE, showWarnings = FALSE)

eval_df <- read_parquet(abs_path("analysis/late_round_eval/data/eval_df.parquet"))
harmonized_all <- read_parquet(abs_path("analysis/late_round_eval/extraction/output/harmonized.parquet"))
models_out <- run_per_position_models(eval_df)

fmt <- function(x, d = 2) ifelse(is.na(x), "—", sprintf(paste0("%.", d, "f"), x))
```

# Executive summary

We define the ground-truth tier from realized production (best PPR FFPG, Y1–Y3 post-draft):

| Production tier | Cutoff (PPR PPG) | Verbal anchor |
|---|---|---|
| **Elite** | ≥ 16.5 | RB1 / WR1 cornerstone |
| **Starter** | 12 – 16.5 | RB2 / WR2 (weekly start) |
| **Flex** | 7 – 12 | FLEX-worthy |
| **Depth** | 3 – 7 | bench |
| **Dart Throw** | < 3 | irrelevant |

We compare JJ's predicted tier (`canonical_tier` from his guide) to this production tier, alongside a baseline tier derived from `lm(best_ffppg ~ age + log(draft_pick))` binned with the same cutoffs.

```{r exec-summary}
tier_eval <- run_tier_eval(eval_df)

tibble::tibble(
  Position = c("WR", "RB"),
  N        = c(tier_eval$WR$his$overall$n, tier_eval$RB$his$overall$n),
  `JJ accuracy`        = c(fmt(tier_eval$WR$his$overall$accuracy, 3),
                            fmt(tier_eval$RB$his$overall$accuracy, 3)),
  `Baseline accuracy`  = c(fmt(tier_eval$WR$baseline$overall$accuracy, 3),
                            fmt(tier_eval$RB$baseline$overall$accuracy, 3)),
  `JJ κ`               = c(fmt(tier_eval$WR$his$overall$quadratic_kappa, 3),
                            fmt(tier_eval$RB$his$overall$quadratic_kappa, 3)),
  `Baseline κ`         = c(fmt(tier_eval$WR$baseline$overall$quadratic_kappa, 3),
                            fmt(tier_eval$RB$baseline$overall$quadratic_kappa, 3)),
  `JJ weighted F1`     = c(fmt(tier_eval$WR$his$overall$weighted_f1, 3),
                            fmt(tier_eval$RB$his$overall$weighted_f1, 3)),
  `Baseline weighted F1` = c(fmt(tier_eval$WR$baseline$overall$weighted_f1, 3),
                              fmt(tier_eval$RB$baseline$overall$weighted_f1, 3)),
  `JJ off-by-one`      = c(fmt(tier_eval$WR$his$overall$off_by_one_rate, 3),
                            fmt(tier_eval$RB$his$overall$off_by_one_rate, 3)),
  `Baseline off-by-one`= c(fmt(tier_eval$WR$baseline$overall$off_by_one_rate, 3),
                            fmt(tier_eval$RB$baseline$overall$off_by_one_rate, 3))
) |> kable(caption = "Headline: JJ's canonical_tier vs baseline lm(best_ffppg ~ age + log(draft_pick)) binned via same cutoffs.")
```

**Headline:** On both positions, the age + log(draft_pick) baseline matches or beats JJ's canonical tier on quadratic kappa and off-by-one rate — confirming the regression result that draft capital captures most of his signal. The interesting story is in the per-tier breakdown below: where exactly he wins and loses.

# Methodology evolution 2022–2025

The guide evolved markedly across the four years studied.

- **2022–2024 (Z-Prospect Model):** Numeric tiers 1–15 (2022, 2023) or 1–12 (2024) drawn directly from the post-draft rookie rankings cheatsheet. No verbal tier semantics beyond ordinal position.
- **2025–2026 (ZAP Model):** Verbal tiers (Elite Producer → Dart Throw, plus a 2026-only Legendary Performer top tier). Each tier carries a production claim (e.g., "Weekly Starter — should provide stability in your starting lineup, think RB2/WR2").

```{r methodology-features}
years <- 2022:2025
feature_rows <- lapply(years, function(y) {
  mp <- abs_path(file.path("analysis/late_round_eval/extraction/output", paste0(y, "_metadata.json")))
  if (!file.exists(mp)) return(NULL)
  m <- jsonlite::fromJSON(mp)
  feats <- m$features_mentioned
  if (is.null(feats)) return(NULL)
  tibble(year = y, feature = as.character(feats))
}) |> bind_rows()

if (nrow(feature_rows) > 0) {
  feature_table <- feature_rows |>
    mutate(present = "✓") |>
    pivot_wider(names_from = year, values_from = present, values_fill = "·")
  kable(feature_table, caption = "Features mentioned in methodology, by year")
}
```

# Tier definitions — his and ours

## How JJ defines each tier

The verbal tier scheme (2025–2026) is the only year-range where JJ publishes explicit definitions. **Each tier is a band of the ZAP Score** (his model output, 0–100 scale; higher = better prospect):

| Canonical bucket | JJ's verbal tier (2025–2026) | Stated ZAP score band | JJ's production claim |
|---|---|---|---|
| Elite | Legendary Performer *(2026 only)* | 90–100 | "Generational talent. Hit at outstanding rates historically. Cornerstone dynasty asset." |
| Elite | Elite Producer | 75–90 | "Best of the best. Legitimate shot at a cornerstone piece." |
| Starter | Weekly Starter | 60–75 | "Stability in your starting lineup. Think RB2/WR2." |
| Flex | Flex Play | 40–60 | "Rarely become league-winners. Can provide usable production." |
| Depth | Benchwarmer | 30–40 | "Perfect for bench insurance." |
| Depth | Waiver Wire Add | 20–30 | "Probably doesn't need to be rostered, but can occasionally be usable." |
| Dart Throw | Dart Throw | 0–20 | "Very little chance for success." |

For **2022–2024**, JJ uses purely **numeric tiers** from his Post-Draft Rookie Rankings Cheatsheet (Tier 1 → Tier 15 in 2022/2023; Tier 1 → Tier 12 in 2024 + an "Unranked" bucket for players he profiled but didn't rank). There is no verbal label or stated FFPG/ZAP cutoff per tier in those years — the "definition" is just rank order in the cheatsheet.

## How WE map his tiers to our canonical 5-tier scheme

To compare across years, we collapse every year's tier nomenclature to **{Elite, Starter, Flex, Depth, Dart Throw}** (canonical 5-tier scheme). The full mapping lives in `extraction/output/tier_map.json`; here's the summary by year:

| Canonical | 2022 (numeric 1–15) | 2023 (Tier 1–15) | 2024 (Tier 1–12 + Unranked) | 2025 / 2026 (verbal) |
|---|---|---|---|---|
| **Elite** | 1 | Tier 1 | Tier 1 | Legendary Performer / Elite Producer |
| **Starter** | 2–3 | Tier 2–3 | Tier 2–3 | Weekly Starter |
| **Flex** | 4–6 | Tier 4–6 | Tier 4–5 | Flex Play |
| **Depth** | 7–9 | Tier 7–9 | Tier 6–8 | Benchwarmer / Waiver Wire Add |
| **Dart Throw** | 10–15 + Z-Prospect Ranked | Tier 10–15 | Tier 9–12 + Unranked | Dart Throw |

Numeric tiers (2022–2024) were collapsed by proportional bucketing: ~top 7% → Elite, next ~15% → Starter, etc. The harmonizer's full rationale is in `tier_map.json` under the `rationale` key.

## Empirical ZAP score distribution by canonical tier (2024–2026)

What ZAP score does each canonical tier actually carry, year by year? (Players without ZAP scores — 2022/2023 cheatsheet-only — are excluded.)

```{r zap-by-tier}
harmonized_all |>
  dplyr::filter(position %in% c("WR","RB"), !is.na(zap_score)) |>
  group_by(guide_year, canonical_tier) |>
  summarise(n = n(),
            mean_zap = round(mean(zap_score), 1),
            min_zap  = round(min(zap_score), 1),
            max_zap  = round(max(zap_score), 1),
            .groups = "drop") |>
  pivot_wider(names_from = canonical_tier,
              values_from = mean_zap,
              id_cols = guide_year) |>
  kable(caption = "Mean ZAP score by canonical tier and year (WR+RB only; 2022/2023 NA)")
```

The 2024 mapping looks compressed near the top because 2024 used many fine-grained numeric tiers; our canonical 5-tier collapse puts everyone Tier 1–5 into Elite/Starter/Flex, all with ZAP scores above 80. 2025/2026 verbal tiers spread across the full ZAP range more evenly.

# Coverage 2022–2026

```{r coverage}
chart_coverage(eval_df, harmonized_all)
```

```{r coverage-table}
harmonized_all |>
  dplyr::filter(position %in% c("WR", "RB")) |>
  count(guide_year, position) |>
  pivot_wider(names_from = position, values_from = n, values_fill = 0) |>
  kable(caption = "WR + RB players per guide year")
```

**Note on 2026:** the 2026 class was extracted and harmonized but is excluded from all eval metrics. Player birthdates aren't yet in nflreadr for these rookies, and even if matched, they have no NFL season data yet.

# Production tier distribution

How many of each FFPG-defined tier did 2022–2025 rookies actually produce?

```{r production-tier-dist}
eval_df |>
  count(position, production_tier) |>
  pivot_wider(names_from = production_tier, values_from = n, values_fill = 0) |>
  kable(caption = "Production tier counts (eval set, 2022-2025)")
```

```{r production-by-tier}
chart_production_by_tier(eval_df)
```

The boxplot still uses JJ's `canonical_tier` on the x-axis — it shows that his Elite/Starter calls do produce higher PPG on average, just with massive variance.

# How JJ's tier compares to the production tier

Below we compare JJ's predicted `canonical_tier` (from his guide) directly to the FFPG-defined `production_tier` (truth). For reference we also compare a baseline that predicts FFPG from `age + log(draft_pick)` and bins the prediction with the same cutoffs.

## Confusion matrices

```{r conf-his-wr, fig.width=8, fig.height=5}
chart_confusion_matrix(tier_eval$WR$his$confusion, "WR — JJ's tier vs production tier")
```

```{r conf-base-wr, fig.width=8, fig.height=5}
chart_confusion_matrix(tier_eval$WR$baseline$confusion, "WR — Baseline (age + log capital) tier vs production tier")
```

```{r conf-his-rb, fig.width=8, fig.height=5}
chart_confusion_matrix(tier_eval$RB$his$confusion, "RB — JJ's tier vs production tier")
```

```{r conf-base-rb, fig.width=8, fig.height=5}
chart_confusion_matrix(tier_eval$RB$baseline$confusion, "RB — Baseline (age + log capital) tier vs production tier")
```

Each row is the true tier (production-derived); each column is the predicted tier. Diagonal = correct, off-diagonal = miss.

## Per-tier precision, recall, F1

```{r per-tier-tables, results='asis'}
for (pos in c("WR", "RB")) {
  cat(sprintf("\n### %s\n\n", pos))
  cat("\n**JJ's predictions:**\n\n")
  print(
    tier_eval[[pos]]$his$per_tier |>
      mutate(across(c(precision, recall, f1), ~ round(.x, 3))) |>
      kable()
  )
  cat("\n**Baseline (age + log capital) predictions:**\n\n")
  print(
    tier_eval[[pos]]$baseline$per_tier |>
      mutate(across(c(precision, recall, f1), ~ round(.x, 3))) |>
      kable()
  )
  cat("\n")
}
```

**Read:** Precision answers "of players he tagged tier X, what fraction actually became tier X." Recall answers "of players who actually were tier X, what fraction did he catch." Both predictors over-call Dart Throw and Depth because most rookies bust — the meaningful question is the Starter/Elite rows.

## Tier-call calibration: do his Elite calls actually become Elite?

The diagonal cells of the confusion matrix above answer this directly. Pulling out the high-stakes tiers (Starter + Elite, where his model is essentially "this guy is fantasy-relevant"):

```{r calibration}
top_calls <- function(eval_obj, pos) {
  conf <- eval_obj$confusion
  tier_levels <- c("Dart Throw", "Depth", "Flex", "Starter", "Elite")
  # for each predicted tier, what fraction of those calls reached that tier or above?
  out <- lapply(tier_levels, function(cls) {
    called <- sum(conf[, cls])
    if (called == 0) return(tibble(position = pos, predicted_tier = cls,
                                    n_called = 0, frac_at_or_above = NA_real_,
                                    frac_exact = NA_real_))
    truth_indices <- which(tier_levels == cls):length(tier_levels)
    at_or_above <- sum(conf[truth_indices, cls])
    exact <- conf[cls, cls]
    tibble(position = pos, predicted_tier = cls, n_called = called,
           frac_at_or_above = at_or_above / called,
           frac_exact = exact / called)
  }) |> bind_rows()
  out
}

bind_rows(
  top_calls(tier_eval$WR$his, "WR") |> mutate(predictor = "JJ"),
  top_calls(tier_eval$WR$baseline, "WR") |> mutate(predictor = "Baseline"),
  top_calls(tier_eval$RB$his, "RB") |> mutate(predictor = "JJ"),
  top_calls(tier_eval$RB$baseline, "RB") |> mutate(predictor = "Baseline")
) |>
  mutate(predicted_tier = factor(predicted_tier,
                                  levels = c("Dart Throw","Depth","Flex","Starter","Elite"),
                                  ordered = TRUE)) |>
  arrange(position, predictor, desc(predicted_tier)) |>
  mutate(across(c(frac_at_or_above, frac_exact),
                ~ ifelse(is.na(.x), "—", scales::percent(.x, accuracy = 1)))) |>
  kable(caption = "Calibration: for each predicted tier, what fraction of calls were exactly correct vs. at-or-above the predicted tier.")
```

# Regression — supplementary

The original regression-style read on the question "does his tier add over draft capital + age?" Same conclusion as the classification: no statistically significant signal.

```{r reg-supp}
tibble::tibble(
  Position = c("WR", "RB"),
  N = c(models_out$WR$regression$metrics$n, models_out$RB$regression$metrics$n),
  `Baseline adj R²` = c(fmt(models_out$WR$regression$metrics$adj_r2_baseline),
                         fmt(models_out$RB$regression$metrics$adj_r2_baseline)),
  `+ canonical_tier adj R²` = c(fmt(models_out$WR$regression$metrics$adj_r2_guide),
                                  fmt(models_out$RB$regression$metrics$adj_r2_guide)),
  `ΔR²` = c(fmt(models_out$WR$regression$metrics$delta_r2),
             fmt(models_out$RB$regression$metrics$delta_r2)),
  `F-test p` = c(fmt(models_out$WR$regression$metrics$f_test_p, 3),
                  fmt(models_out$RB$regression$metrics$f_test_p, 3))
) |> kable()
```

# Slices — where (if anywhere) does he add value?

```{r slice-heatmap}
chart_slice_heatmap(eval_df)
```

```{r slice-table}
eval_df |>
  group_by(position, draft_round, p5_flag) |>
  summarise(n = n(),
            mean_ffppg = mean(best_ffppg, na.rm = TRUE),
            starter_plus_rate = mean(production_tier >= "Starter"),
            .groups = "drop") |>
  mutate(mean_ffppg = round(mean_ffppg, 2),
         starter_plus_rate = scales::percent(starter_plus_rate, accuracy = 1)) |>
  kable(caption = "Production by position × draft round × P5. starter_plus_rate = fraction that reached production_tier Starter or Elite (best FFPG ≥ 12).")
```

# Sleeper detection — when JJ disagrees with draft capital

**Operational definition:** a JJ sleeper is a player where JJ's ZAP score (0–100) is higher than the draft-capital-and-age baseline score (also 0–100, percentile-ranked within position). The baseline is `lm(best_ffppg ~ age + log(draft_pick))` predictions, ranked.

```{r score-table-build, results='asis'}
scored <- build_score_table(eval_df)
cat(sprintf("\n**JJ score** is on a 0-100 scale: ZAP score directly for 2024-2026 (where JJ publishes it per-player), and a tier midpoint for 2022-2023 (where the cheatsheet only gives ordinal tier; we use JJ's own stated 2026 ZAP-band midpoints — Dart Throw=10, Depth=30, Flex=50, Starter=67.5, Elite=87.5). **Baseline score** is the percentile rank within position of `lm(best_ffppg ~ age + log(draft_pick))` predictions.\n\n"))
cat(sprintf("Coverage: **%d total players** in the score table — **%d** with native ZAP (2024-25), **%d** with tier-midpoint fallback (2022-23).\n\n",
            nrow(scored), sum(!is.na(scored$zap_score)), sum(is.na(scored$zap_score))))
```

## How often does JJ disagree with draft capital?

```{r jj-vs-baseline-counts}
scored |>
  dplyr::mutate(direction = dplyr::case_when(
    jj_score > baseline_score + 10 ~ "JJ more bullish (+10 or more)",
    jj_score < baseline_score - 10 ~ "JJ more bearish (-10 or worse)",
    TRUE                           ~ "Roughly aligned (within +-10)"
  )) |>
  count(position, direction) |>
  pivot_wider(names_from = direction, values_from = n, values_fill = 0) |>
  kable(caption = "Counts of (JJ score) vs (baseline percentile score) disagreement, 2022-2025 full eval set")
```

## The scatter — score-vs-score, colored by actual production tier

Each point is a 2022–2025 rookie. The diagonal is "JJ agrees with draft capital." Points above = JJ-bumps; below = JJ-fades. Color = actual production tier. Triangles = day-3/UDFA, circles = early-round. Top 5 biggest disagreements each way (per position) are name-labeled.

```{r score-scatter, fig.height=7, fig.width=11}
chart_score_scatter(scored, label_top_n = 5)
```

## Outcomes by JJ-bump magnitude

If JJ has any independent signal, players he scores meaningfully higher than the baseline should hit Starter+ at higher rates than players where he agrees with the baseline (or fades).

```{r bump-outcomes-full}
bump_full <- jj_bump_analysis(scored)
bump_full$summary |>
  mutate(across(c(mean_ffppg), ~ round(.x, 2)),
         across(c(starter_plus_rate, elite_rate, bust_rate),
                ~ scales::percent(.x, accuracy = 1))) |>
  kable(caption = "Outcomes by JJ-vs-baseline gap bucket — full eval set (2024-2025)")
```

```{r bump-chart-full, fig.height=4.5}
chart_bump_outcomes(bump_full$summary)
```

## Late-round only (day-3 + UDFA)

```{r bump-late}
bump_late <- jj_bump_analysis(scored, late_round_only = TRUE)
bump_late$summary |>
  mutate(across(c(mean_ffppg), ~ round(.x, 2)),
         across(c(starter_plus_rate, elite_rate, bust_rate),
                ~ scales::percent(.x, accuracy = 1))) |>
  kable(caption = "Outcomes by JJ-vs-baseline gap bucket — late-round only (day-3 + UDFA)")
```

## Named JJ-bump cases (biggest disagreements where JJ was more bullish)

```{r jj-bumps-named}
bump_full$detail |>
  dplyr::filter(bump > 0) |>
  arrange(desc(bump)) |>
  head(15) |>
  dplyr::select(name, position, guide_year, draft_round, baseline_score, jj_score, bump,
                production_tier, best_ffppg) |>
  mutate(across(c(baseline_score, jj_score, bump), ~ round(.x, 1)),
         best_ffppg = round(best_ffppg, 1)) |>
  kable(caption = "Top 15 JJ-bump players — biggest positive disagreement vs baseline. production_tier and best_ffppg show what actually happened.")
```

## Named JJ-fade cases (biggest disagreements where JJ was more bearish)

```{r jj-fades-named}
bump_full$detail |>
  dplyr::filter(bump < 0) |>
  arrange(bump) |>
  head(15) |>
  dplyr::select(name, position, guide_year, draft_round, baseline_score, jj_score, bump,
                production_tier, best_ffppg) |>
  mutate(across(c(baseline_score, jj_score, bump), ~ round(.x, 1)),
         best_ffppg = round(best_ffppg, 1)) |>
  kable(caption = "Top 15 JJ-fade players — biggest negative disagreement vs baseline.")
```

# Ranking quality — for dynasty drafting

If you're using one of these rankings to draft (1.01, 1.02, 1.03...), how well does each one order players against actual production? Three views.

## Spearman rank correlation with actual best_ffppg

```{r spearman}
rank_out <- ranking_quality(scored)
tibble::tibble(
  Position = c("WR", "RB"),
  N = c(rank_out$WR$n, rank_out$RB$n),
  `Spearman: JJ vs actual`       = c(fmt(rank_out$WR$spearman_jj, 3),
                                      fmt(rank_out$RB$spearman_jj, 3)),
  `Spearman: Baseline vs actual` = c(fmt(rank_out$WR$spearman_baseline, 3),
                                      fmt(rank_out$RB$spearman_baseline, 3))
) |> kable(caption = "Spearman rank correlation of each score with actual best_ffppg.")
```

## Top-K precision — of the top K by each ranking, what fraction were truly top K?

```{r top-k}
dplyr::bind_rows(
  rank_out$WR$top_k |> mutate(position = "WR"),
  rank_out$RB$top_k |> mutate(position = "RB")
) |>
  dplyr::select(position, k, jj_precision, baseline_precision) |>
  mutate(across(c(jj_precision, baseline_precision), ~ scales::percent(.x, accuracy = 1))) |>
  kable(caption = "Of the top K players each ranker selected, fraction that were actually top K by best_ffppg.")
```

## Cumulative production curve

If you drafted strictly in each ranker's order, this is the total `best_ffppg` you would have accumulated as you went down the list. Steeper = better ranker (picks the productive players first). The dashed line is the oracle upper bound (ranking by actual production).

```{r ranking-lift, fig.height=4.5, fig.width=11}
chart_ranking_lift(rank_out)
```

## Pearson correlation with continuous best_ffppg

```{r score-correlations}
score_correlations(scored) |>
  mutate(across(where(is.numeric), ~ round(.x, 3))) |>
  kable(caption = "Pearson correlation of each score with continuous best_ffppg. pearson_zap_only_2024_25 restricts to ZAP-available years for a like-for-like ZAP-vs-baseline comparison.")
```

# Year-by-year scorecard

```{r yearly, results='asis'}
for (y in 2022:2025) {
  cat(sprintf("\n## %d class\n\n", y))
  print(
    eval_df |>
      dplyr::filter(guide_year == y) |>
      arrange(desc(as.integer(canonical_tier)), draft_pick) |>
      dplyr::select(name, position, canonical_tier, draft_round, best_ffppg) |>
      mutate(best_ffppg = round(best_ffppg, 2)) |>
      kable()
  )
}
```

# Sensitivity appendix

```{r sensitivity}
wr_df <- eval_df |> dplyr::filter(position == "WR")
rb_df <- eval_df |> dplyr::filter(position == "RB")

wr_yearfe_b <- lm(best_ffppg ~ age + log(draft_pick) + factor(guide_year), data = wr_df)
wr_yearfe_g <- lm(best_ffppg ~ age + log(draft_pick) + factor(guide_year) + canonical_tier, data = wr_df)
rb_yearfe_b <- lm(best_ffppg ~ age + log(draft_pick) + factor(guide_year), data = rb_df)
rb_yearfe_g <- lm(best_ffppg ~ age + log(draft_pick) + factor(guide_year) + canonical_tier, data = rb_df)

tibble::tibble(
  Position = c("WR", "RB"),
  `ΔR² (no year FE)` = c(fmt(models_out$WR$regression$metrics$delta_r2),
                          fmt(models_out$RB$regression$metrics$delta_r2)),
  `ΔR² (with year FE)` = c(
    fmt(summary(wr_yearfe_g)$adj.r.squared - summary(wr_yearfe_b)$adj.r.squared),
    fmt(summary(rb_yearfe_g)$adj.r.squared - summary(rb_yearfe_b)$adj.r.squared)
  )
) |> kable(caption = "Sensitivity: effect of including year fixed effects")
```

### Match funnel summary

```{r match-funnel}
matches <- read_parquet(abs_path("analysis/late_round_eval/extraction/output/matches.parquet"))
matches |> count(match_stage) |> kable()
```

---

*Built with the late round prospect guide evaluation pipeline. Spec: `docs/superpowers/specs/2026-05-16-late-round-eval-design.md`. Plan: `docs/superpowers/plans/2026-05-16-late-round-eval.md`.*
