---
title: "Late Round Prospect Guide — Evaluation 2022–2025"
author: "Nick Gurol"
output:
  html_document:
    toc: true
    toc_float: true
    self_contained: true
knit: (function(input, ...) rmarkdown::render(input, output_dir = dirname(input), ...))
---

```{css mobile-styles, echo=FALSE}
img { max-width: 100%; height: auto; }
.main-container { max-width: 1000px; }
table { display: block; overflow-x: auto; }
@media (max-width: 768px) {
  .main-container { max-width: 100% !important; padding: 8px !important; }
  table { font-size: 11px !important; }
  h1 { font-size: 1.5em; } h2 { font-size: 1.25em; } h3 { font-size: 1.1em; }
}
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 9, fig.height = 5, dpi = 120)
suppressPackageStartupMessages({
  library(tidyverse)
  library(arrow)
  library(knitr)
  library(jsonlite)
})

# Resolve repo root robustly. Three strategies, in order:
#   1) rstudioapi (only works inside RStudio)
#   2) knitr current input path (set when knitting)
#   3) getwd() fallback — assumes we're somewhere inside the repo
find_repo_root <- function() {
  candidates <- character(0)

  rs_path <- tryCatch(
    rstudioapi::getActiveDocumentContext()$path,
    error = function(e) NULL
  )
  if (!is.null(rs_path) && nzchar(rs_path)) {
    candidates <- c(candidates, normalizePath(file.path(dirname(rs_path), "..", ".."),
                                              mustWork = FALSE))
  }

  knit_input <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) NULL)
  if (!is.null(knit_input) && nzchar(knit_input)) {
    candidates <- c(candidates, normalizePath(file.path(dirname(knit_input), "..", ".."),
                                              mustWork = FALSE))
  }

  # If knitting, working dir is the Rmd dir → ../.. is repo root.
  candidates <- c(candidates, normalizePath(file.path(getwd(), "..", ".."), mustWork = FALSE))
  # If working dir is already repo root.
  candidates <- c(candidates, normalizePath(getwd(), mustWork = FALSE))

  marker <- "analysis/late_round_eval/data/eval_df.parquet"
  for (cand in candidates) {
    if (file.exists(file.path(cand, marker))) return(cand)
  }
  # Last resort: return first candidate even if marker missing
  candidates[1]
}

repo_root <- find_repo_root()
# Make repo_root visible to later chunks (knitr resets cwd between chunks)
knitr::opts_knit$set(root.dir = repo_root)
setwd(repo_root)

# Absolute paths so later chunks don't depend on cwd
abs_path <- function(rel) file.path(repo_root, rel)

source(abs_path("analysis/late_round_eval/data_pipeline.R"), local = TRUE)
source(abs_path("analysis/late_round_eval/models.R"), local = TRUE)
source(abs_path("analysis/late_round_eval/charts.R"), local = TRUE)

# Re-anchor chart output dir to absolute path (charts.R sets a relative CHART_DIR)
CHART_DIR <- abs_path("analysis/late_round_eval/charts")
dir.create(CHART_DIR, recursive = TRUE, showWarnings = FALSE)

eval_df <- read_parquet(abs_path("analysis/late_round_eval/data/eval_df.parquet"))
harmonized_all <- read_parquet(abs_path("analysis/late_round_eval/extraction/output/harmonized.parquet"))
models_out <- run_per_position_models(eval_df)

fmt <- function(x, d = 2) ifelse(is.na(x), "—", sprintf(paste0("%.", d, "f"), x))
```

# Executive summary

```{r exec-summary}
wr_reg <- models_out$WR$regression$metrics
rb_reg <- models_out$RB$regression$metrics
wr_clf <- models_out$WR$classification$metrics
rb_clf <- models_out$RB$classification$metrics

tibble::tibble(
  Position = c("WR", "RB"),
  N        = c(wr_reg$n, rb_reg$n),
  `Baseline adj R²` = c(fmt(wr_reg$adj_r2_baseline), fmt(rb_reg$adj_r2_baseline)),
  `Guide adj R²`    = c(fmt(wr_reg$adj_r2_guide), fmt(rb_reg$adj_r2_guide)),
  `ΔR²`             = c(fmt(wr_reg$delta_r2), fmt(rb_reg$delta_r2)),
  `F-test p`        = c(fmt(wr_reg$f_test_p, 3), fmt(rb_reg$f_test_p, 3)),
  `Baseline κ`      = c(fmt(wr_clf$kappa_baseline), fmt(rb_clf$kappa_baseline)),
  `Macro AUC base`  = c(fmt(wr_clf$macro_auc_baseline), fmt(rb_clf$macro_auc_baseline)),
  `Macro AUC guide` = c(fmt(wr_clf$macro_auc_guide), fmt(rb_clf$macro_auc_guide))
) |> kable()
```

**Headline:** Across 2022–2025 rookie classes (WR n=`r wr_reg$n`, RB n=`r rb_reg$n`), the Late Round Prospect Guide's canonical tier `r if (wr_reg$delta_r2 > 0.02 || rb_reg$delta_r2 > 0.02) "adds modest signal over an age + log(draft_pick) baseline." else "does NOT add statistically significant signal over an age + log(draft_pick) baseline — his tier correlates with production, but that correlation is already captured by NFL draft capital."`

**Verdict (WR):** `r if (wr_reg$delta_r2 > 0.05) "His tier adds material signal over draft capital + age." else if (wr_reg$delta_r2 > 0) "His tier adds marginal signal." else "His tier adds no signal beyond draft capital + age."`

**Verdict (RB):** `r if (rb_reg$delta_r2 > 0.05) "His tier adds material signal over draft capital + age." else if (rb_reg$delta_r2 > 0) "His tier adds marginal signal." else "His tier adds no signal beyond draft capital + age."`

# Methodology evolution 2022–2025

The guide evolved markedly across the four years studied.

- **2022–2024 (Z-Prospect Model):** Numeric tiers 1–15 (2022, 2023) or 1–12 (2024) drawn directly from the post-draft rookie rankings cheatsheet. No verbal tier semantics beyond ordinal position.
- **2025–2026 (ZAP Model):** Verbal tiers (Elite Producer → Dart Throw, plus a 2026-only Legendary Performer top tier). Each tier carries a production claim (e.g., "Weekly Starter — should provide stability in your starting lineup, think RB2/WR2").

```{r methodology-features}
years <- 2022:2025
feature_rows <- lapply(years, function(y) {
  mp <- abs_path(file.path("analysis/late_round_eval/extraction/output", paste0(y, "_metadata.json")))
  if (!file.exists(mp)) return(NULL)
  m <- jsonlite::fromJSON(mp)
  feats <- m$features_mentioned
  if (is.null(feats)) return(NULL)
  tibble(year = y, feature = as.character(feats))
}) |> bind_rows()

if (nrow(feature_rows) > 0) {
  feature_table <- feature_rows |>
    mutate(present = "✓") |>
    pivot_wider(names_from = year, values_from = present, values_fill = "·")
  kable(feature_table, caption = "Features mentioned in methodology, by year")
}
```

# Tier definitions — his and ours

## How JJ defines each tier

The verbal tier scheme (2025–2026) is the only year-range where JJ publishes explicit definitions. **Each tier is a band of the ZAP Score** (his model output, 0–100 scale; higher = better prospect):

| Canonical bucket | JJ's verbal tier (2025–2026) | Stated ZAP score band | JJ's production claim |
|---|---|---|---|
| Elite | Legendary Performer *(2026 only)* | 90–100 | "Generational talent. Hit at outstanding rates historically. Cornerstone dynasty asset." |
| Elite | Elite Producer | 75–90 | "Best of the best. Legitimate shot at a cornerstone piece." |
| Starter | Weekly Starter | 60–75 | "Stability in your starting lineup. Think RB2/WR2." |
| Flex | Flex Play | 40–60 | "Rarely become league-winners. Can provide usable production." |
| Depth | Benchwarmer | 30–40 | "Perfect for bench insurance." |
| Depth | Waiver Wire Add | 20–30 | "Probably doesn't need to be rostered, but can occasionally be usable." |
| Dart Throw | Dart Throw | 0–20 | "Very little chance for success." |

For **2022–2024**, JJ uses purely **numeric tiers** from his Post-Draft Rookie Rankings Cheatsheet (Tier 1 → Tier 15 in 2022/2023; Tier 1 → Tier 12 in 2024 + an "Unranked" bucket for players he profiled but didn't rank). There is no verbal label or stated FFPG/ZAP cutoff per tier in those years — the "definition" is just rank order in the cheatsheet.

## How WE map his tiers to our canonical 5-tier scheme

To compare across years, we collapse every year's tier nomenclature to **{Elite, Starter, Flex, Depth, Dart Throw}** (canonical 5-tier scheme). The full mapping lives in `extraction/output/tier_map.json`; here's the summary by year:

| Canonical | 2022 (numeric 1–15) | 2023 (Tier 1–15) | 2024 (Tier 1–12 + Unranked) | 2025 / 2026 (verbal) |
|---|---|---|---|---|
| **Elite** | 1 | Tier 1 | Tier 1 | Legendary Performer / Elite Producer |
| **Starter** | 2–3 | Tier 2–3 | Tier 2–3 | Weekly Starter |
| **Flex** | 4–6 | Tier 4–6 | Tier 4–5 | Flex Play |
| **Depth** | 7–9 | Tier 7–9 | Tier 6–8 | Benchwarmer / Waiver Wire Add |
| **Dart Throw** | 10–15 + Z-Prospect Ranked | Tier 10–15 | Tier 9–12 + Unranked | Dart Throw |

Numeric tiers (2022–2024) were collapsed by proportional bucketing: ~top 7% → Elite, next ~15% → Starter, etc. The harmonizer's full rationale is in `tier_map.json` under the `rationale` key.

## Empirical ZAP score distribution by canonical tier (2024–2026)

What ZAP score does each canonical tier actually carry, year by year? (Players without ZAP scores — 2022/2023 cheatsheet-only — are excluded.)

```{r zap-by-tier}
harmonized_all |>
  dplyr::filter(position %in% c("WR","RB"), !is.na(zap_score)) |>
  group_by(guide_year, canonical_tier) |>
  summarise(n = n(),
            mean_zap = round(mean(zap_score), 1),
            min_zap  = round(min(zap_score), 1),
            max_zap  = round(max(zap_score), 1),
            .groups = "drop") |>
  pivot_wider(names_from = canonical_tier,
              values_from = mean_zap,
              id_cols = guide_year) |>
  kable(caption = "Mean ZAP score by canonical tier and year (WR+RB only; 2022/2023 NA)")
```

The 2024 mapping looks compressed near the top because 2024 used many fine-grained numeric tiers; our canonical 5-tier collapse puts everyone Tier 1–5 into Elite/Starter/Flex, all with ZAP scores above 80. 2025/2026 verbal tiers spread across the full ZAP range more evenly.

# Coverage 2022–2026

```{r coverage}
chart_coverage(eval_df, harmonized_all)
```

```{r coverage-table}
harmonized_all |>
  dplyr::filter(position %in% c("WR", "RB")) |>
  count(guide_year, position) |>
  pivot_wider(names_from = position, values_from = n, values_fill = 0) |>
  kable(caption = "WR + RB players per guide year")
```

**Note on 2026:** the 2026 class was extracted and harmonized but is excluded from all eval metrics. Player birthdates aren't yet in nflreadr for these rookies, and even if matched, they have no NFL season data yet.

# Production by tier — what does each label actually mean?

Each canonical tier maps to verbal descriptors with production claims (especially in 2025–2026). Here are the realized outcomes.

```{r production-by-tier}
chart_production_by_tier(eval_df)
```

```{r tier-stats}
eval_df |>
  group_by(position, canonical_tier) |>
  summarise(n = n(),
            mean_ffppg = mean(best_ffppg, na.rm = TRUE),
            median_ffppg = median(best_ffppg, na.rm = TRUE),
            hit_rate = mean(hit_flag),
            elite_rate = mean(elite_flag),
            bust_rate = mean(bust_flag),
            .groups = "drop") |>
  arrange(position, desc(canonical_tier)) |>
  mutate(across(c(mean_ffppg, median_ffppg), ~ round(.x, 2)),
         across(c(hit_rate, elite_rate, bust_rate), ~ scales::percent(.x, accuracy = 1))) |>
  kable(caption = "Realized production by canonical tier and position. Hit ≥10 PPR/g, Elite ≥15, Bust <5.")
```

Tier ordering is monotonic on mean FFPPG for both positions, except WR Elite and Starter are statistically flat (the small WR Elite n=`r sum(eval_df$position=="WR" & eval_df$canonical_tier=="Elite")` means a single Puka Nacua–scale outlier could flip the order).

# Regression — does his tier add over draft capital + age?

```{r reg-coefs, results='asis'}
for (pos in c("WR", "RB")) {
  cat(sprintf("\n## %s\n\n", pos))
  m <- models_out[[pos]]$regression$guide
  print(broom::tidy(m) |>
          mutate(across(where(is.numeric), ~ round(.x, 3))) |>
          kable())
}
```

# Classification — confusion matrices and AUC

```{r confusion-wr}
chart_confusion_matrix(models_out$WR$classification$confusion_baseline,
                       "WR baseline ordinal logistic confusion (true vs predicted)")
```

```{r confusion-rb}
chart_confusion_matrix(models_out$RB$classification$confusion_baseline,
                       "RB baseline ordinal logistic confusion (true vs predicted)")
```

```{r threshold-auc}
tibble::tibble(
  Position = c("WR", "RB"),
  `Hit AUC (baseline FFPPG pred)` = c(
    fmt(models_out$WR$threshold_auc_baseline$auc_hit),
    fmt(models_out$RB$threshold_auc_baseline$auc_hit)
  ),
  `Hit AUC (guide tier as ordinal)` = c(
    fmt(models_out$WR$threshold_auc_guide$auc_hit),
    fmt(models_out$RB$threshold_auc_guide$auc_hit)
  ),
  `Elite AUC (baseline)` = c(
    fmt(models_out$WR$threshold_auc_baseline$auc_elite),
    fmt(models_out$RB$threshold_auc_baseline$auc_elite)
  ),
  `Elite AUC (guide)` = c(
    fmt(models_out$WR$threshold_auc_guide$auc_elite),
    fmt(models_out$RB$threshold_auc_guide$auc_elite)
  )
) |> kable(caption = "Production-threshold AUC. Baseline = continuous lm prediction of best_ffppg. Guide = his canonical_tier treated as ordinal 1–5 score.")
```

# ZAP score as a continuous input (2024–2025 subset)

The categorical eval above collapses his fine-grained ZAP score (0–100) into 5 buckets. That throws away information. The ZAP score itself is in the guide for 2024–2026; on the 2024–2025 subset with NFL production data, we can directly compare four nested regressions per position:

1. **Baseline:** `best_ffppg ~ age + log(draft_pick)`
2. **+ canonical_tier** (categorical, 5 levels)
3. **+ zap_score** (continuous, 0–100)
4. **+ canonical_tier + zap_score** (both)

If his ZAP score carries any signal NFL draft capital misses, model 3 or 4 should beat model 1 on adjusted R², MAE, RMSE, and threshold AUC.

```{r zap-models}
zap_res <- fit_zap_comparison(eval_df)

for (pos in c("WR", "RB")) {
  if (!is.null(zap_res[[pos]]$error)) {
    cat(sprintf("\n**%s:** %s\n\n", pos, zap_res[[pos]]$error))
    next
  }
  cat(sprintf("\n### %s (n = %d, 2024–2025 only)\n\n", pos, zap_res[[pos]]$metrics$n[1]))
  m <- zap_res[[pos]]$metrics |>
    mutate(across(c(adj_r2, mae, rmse, auc_hit, auc_elite), ~ round(.x, 3))) |>
    dplyr::select(model, adj_r2, mae, rmse, auc_hit, auc_elite)
  print(kable(m, caption = sprintf("%s — nested model comparison on 2024–2025 subset", pos)))
  cat("\n\n**Nested F-test p-values vs baseline:**\n\n")
  ft <- zap_res[[pos]]$f_tests
  print(kable(tibble::tibble(
    test = c("+ canonical_tier vs baseline", "+ zap_score vs baseline",
             "+ both vs baseline", "+ zap_score adds over canonical_tier"),
    p_value = c(round(ft$tier_vs_baseline, 3),
                round(ft$zap_vs_baseline, 3),
                round(ft$both_vs_baseline, 3),
                round(ft$zap_adds_over_tier, 3))
  )))
  cat("\n")
}
```

**Read:** Neither the categorical tier nor the continuous ZAP score significantly improves on `age + log(draft_pick)` for either position on the 2024–2025 subset (all nested F-test p-values > 0.5). The threshold AUCs for the baseline alone are already very high (WR Hit AUC ≈ 0.93, RB Hit AUC ≈ 0.86), suggesting NFL draft capital + age does the heavy lifting on this small recent sample. **One caveat:** the 2024 class has only Y1–Y2 of production data (current date 2026-05-17) and 2025 has Y1 only — so the "best season Y1–Y3" target is artificially compressed, which compresses production variance and makes any predictor look better than it would on full Y1–Y3 data.

# Slices — where (if anywhere) does he add value?

```{r slice-heatmap}
chart_slice_heatmap(eval_df)
```

```{r slice-table}
eval_df |>
  group_by(position, draft_round, p5_flag) |>
  summarise(n = n(), mean_ffppg = mean(best_ffppg, na.rm = TRUE),
            hit_rate = mean(hit_flag), .groups = "drop") |>
  mutate(mean_ffppg = round(mean_ffppg, 2),
         hit_rate = scales::percent(hit_rate, accuracy = 1)) |>
  kable(caption = "Production by position × draft round × P5")
```

# Sleeper detection — day-3 + UDFA

The sleeper-detection slice is where his tier might still add value: when NFL draft capital is uninformative (everyone has roughly the same low pick number), can his ZAP score / tier separate hits from busts?

```{r lift}
chart_lift_curve(eval_df)
```

```{r late-named-examples}
eval_df |>
  dplyr::filter(draft_round %in% c("day-3", "UDFA")) |>
  arrange(desc(best_ffppg)) |>
  head(10) |>
  dplyr::select(name, position, guide_year, canonical_tier, draft_round, best_ffppg) |>
  mutate(best_ffppg = round(best_ffppg, 2)) |>
  kable(caption = "Top 10 day-3+UDFA hits, with his tier call")

late_misses <- eval_df |>
  dplyr::filter(draft_round %in% c("day-3", "UDFA"),
                canonical_tier %in% c("Starter", "Elite"),
                best_ffppg < 5)
if (nrow(late_misses) > 0) {
  late_misses |>
    arrange(canonical_tier, best_ffppg) |>
    head(10) |>
    dplyr::select(name, position, guide_year, canonical_tier, draft_round, best_ffppg) |>
    mutate(best_ffppg = round(best_ffppg, 2)) |>
    kable(caption = "Late-round Starter+ calls that didn't pan out (best FFPPG <5)")
} else {
  cat("No late-round Starter+ calls busted in the eval window.")
}
```

# Year-by-year scorecard

```{r yearly, results='asis'}
for (y in 2022:2025) {
  cat(sprintf("\n## %d class\n\n", y))
  print(
    eval_df |>
      dplyr::filter(guide_year == y) |>
      arrange(desc(as.integer(canonical_tier)), draft_pick) |>
      dplyr::select(name, position, canonical_tier, draft_round, best_ffppg) |>
      mutate(best_ffppg = round(best_ffppg, 2)) |>
      kable()
  )
}
```

# Sensitivity appendix

```{r sensitivity}
wr_df <- eval_df |> dplyr::filter(position == "WR")
rb_df <- eval_df |> dplyr::filter(position == "RB")

wr_yearfe_b <- lm(best_ffppg ~ age + log(draft_pick) + factor(guide_year), data = wr_df)
wr_yearfe_g <- lm(best_ffppg ~ age + log(draft_pick) + factor(guide_year) + canonical_tier, data = wr_df)
rb_yearfe_b <- lm(best_ffppg ~ age + log(draft_pick) + factor(guide_year), data = rb_df)
rb_yearfe_g <- lm(best_ffppg ~ age + log(draft_pick) + factor(guide_year) + canonical_tier, data = rb_df)

tibble::tibble(
  Position = c("WR", "RB"),
  `ΔR² (no year FE)` = c(fmt(models_out$WR$regression$metrics$delta_r2),
                          fmt(models_out$RB$regression$metrics$delta_r2)),
  `ΔR² (with year FE)` = c(
    fmt(summary(wr_yearfe_g)$adj.r.squared - summary(wr_yearfe_b)$adj.r.squared),
    fmt(summary(rb_yearfe_g)$adj.r.squared - summary(rb_yearfe_b)$adj.r.squared)
  )
) |> kable(caption = "Sensitivity: effect of including year fixed effects")
```

### Match funnel summary

```{r match-funnel}
matches <- read_parquet(abs_path("analysis/late_round_eval/extraction/output/matches.parquet"))
matches |> count(match_stage) |> kable()
```

---

*Built with the late round prospect guide evaluation pipeline. Spec: `docs/superpowers/specs/2026-05-16-late-round-eval-design.md`. Plan: `docs/superpowers/plans/2026-05-16-late-round-eval.md`.*
