---
title: "NFL Draft Quasi-Sharpe Ratio Analysis"
subtitle: "How much is a draft pick really worth, by position?"
output:
  github_document:
    html_preview: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.width = 10, fig.height = 7, dpi = 300,
  fig.path = "output/charts/"
)
library(tidyverse)
library(arrow)
library(gt)
library(scales)
library(ggridges)

pos_colors <- c(
  "QB" = "#E41A1C", "RB" = "#377EB8", "WR" = "#4DAF4A", "TE" = "#984EA3",
  "OT" = "#FF7F00", "IOL" = "#FFFF33", "EDGE" = "#A65628", "DL" = "#F781BF",
  "LB" = "#999999", "CB" = "#66C2A5", "S" = "#FC8D62"
)

tier_order <- c("Top 10", "Late 1st", "Day 2", "Day 3 Early", "Day 3 Late")
```

```{r load-data}
df <- read_parquet("data/draft_sharpe_analysis.parquet")
sharpe <- read_csv("output/sharpe_ratios.csv", show_col_types = FALSE)
fa_replacement <- read_csv("output/fa_replacement.csv", show_col_types = FALSE)
player_lookup <- read_csv("output/player_lookup.csv", show_col_types = FALSE)

sharpe <- sharpe %>% mutate(tier = factor(tier, levels = tier_order))
df <- df %>% mutate(tier = factor(tier, levels = tier_order))
player_lookup <- player_lookup %>% mutate(tier = factor(tier, levels = tier_order))
```

## The Setup

Traditional draft pick valuation treats picks as interchangeable assets -- a "top 10 pick" has
one value regardless of whether you're drafting a quarterback or a safety. But the return on
investment varies enormously by position. A quasi-Sharpe ratio lets us quantify this.

**The formula:**

$$\text{Sharpe}_{\text{linear}} = \frac{\overline{V}_{\text{2nd contract}} - V_{\text{FA replacement}}}{\sigma_{\text{2nd contract}}}$$

| Component | What it measures |
|---|---|
| Mean second-contract value (% of cap) | The average return on a drafted player at that position and tier |
| FA replacement cost (median cap %) | What you'd pay on the open market for a replacement-level player |
| Standard deviation of second-contract value | The variance (risk) in outcomes |

A **hit** is defined as a player whose average snap percentage exceeds the position-specific
threshold (derived from the median snap % of all players at that position who played 3+ seasons).

The key insight is **nonlinearity**: the distribution of second-contract values is heavily
right-skewed. A few elite players earn massive contracts while most earn close to replacement
cost. This means the linear Sharpe ratio understates the value of positions where elite outcomes
are disproportionately valuable. The elite-weighted variant accounts for this:

$$\text{Sharpe}_{\text{elite}} = \frac{P(\text{elite}) - P(\text{elite})_{\text{baseline}}}{\sigma_{\text{2nd contract}}}$$

where "elite" is defined as a second-contract value in the 90th percentile.

```{r elite-threshold}
elite_threshold <- quantile(df$second_apy_cap_pct[df$second_apy_cap_pct > 0], 0.90)
```

## Hit Rates by Position and Tier

How often does a draft pick become a starter-caliber player? This heatmap shows hit rates
(players exceeding the snap % threshold) by position group and draft tier.

```{r hit-rate-heatmap, fig.height=8}
hit_rates <- sharpe %>%
  select(pos_group, tier, hit_rate, n)

ggplot(hit_rates, aes(x = tier, y = reorder(pos_group, hit_rate), fill = hit_rate)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = paste0(round(hit_rate * 100), "%\n(n=", n, ")")),
            size = 3.2, color = "white", fontface = "bold") +
  scale_fill_gradient2(
    low = "#c62828", mid = "#f9a825", high = "#2e7d32",
    midpoint = 0.5, limits = c(0, 1),
    labels = percent_format()
  ) +
  labs(
    title = "Draft Hit Rates by Position Group and Tier",
    subtitle = "Hit = average snap % exceeds position-specific threshold",
    x = NULL, y = NULL, fill = "Hit Rate"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    panel.grid = element_blank()
  )
```

```{r hit-rate-table}
hit_rates %>%
  mutate(hit_rate = percent(hit_rate, accuracy = 0.1)) %>%
  select(-n) %>%
  pivot_wider(names_from = tier, values_from = hit_rate) %>%
  gt() %>%
  tab_header(title = "Hit Rates by Position Group and Draft Tier") %>%
  cols_label(pos_group = "Position")
```

## The Nonlinearity Story

Second-contract values are not normally distributed. The ridgeline plot below shows how
outcomes cluster near zero for most players, with long right tails for positions like QB and EDGE.

```{r ridgeline-plot, fig.height=8}
df_contracts <- df %>%
  filter(second_apy_cap_pct > 0)

ggplot(df_contracts, aes(x = second_apy_cap_pct, y = reorder(pos_group, second_apy_cap_pct, FUN = median), fill = pos_group)) +
  geom_density_ridges(alpha = 0.7, scale = 1.5, rel_min_height = 0.01) +
  geom_vline(xintercept = elite_threshold, linetype = "dashed", color = "red", linewidth = 0.7) +
  annotate("text", x = elite_threshold + 0.005, y = 1, label = paste0("Elite threshold: ", percent(elite_threshold, accuracy = 0.1)),
           hjust = 0, size = 3.5, color = "red") +
  scale_fill_manual(values = pos_colors) +
  scale_x_continuous(labels = percent_format()) +
  labs(
    title = "Distribution of Second-Contract Values by Position",
    subtitle = "Only players who received a second contract | Dashed line = 90th percentile (elite)",
    x = "Second Contract APY (% of Salary Cap)", y = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
```

```{r elite-probability, fig.height=6}
key_positions <- c("QB", "RB", "WR", "TE", "EDGE", "OT")

elite_by_tier <- sharpe %>%
  filter(pos_group %in% key_positions)

ggplot(elite_by_tier, aes(x = tier, y = elite_prob, color = pos_group, group = pos_group)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_color_manual(values = pos_colors) +
  scale_y_continuous(labels = percent_format()) +
  labs(
    title = "Probability of an Elite Second Contract by Position and Tier",
    subtitle = "Elite = second-contract APY in the 90th percentile | Key positions only",
    x = NULL, y = "P(Elite)", color = "Position"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Free Agency Replacement Cost

The denominator concept: what does it cost to replace a player at each position via free agency?
This baseline is subtracted from draft pick returns to isolate the excess value of drafting.

```{r fa-replacement-bar, fig.height=6}
ggplot(fa_replacement, aes(x = reorder(pos_group, fa_median_apy_cap_pct), y = fa_median_apy_cap_pct, fill = pos_group)) +
  geom_col(width = 0.7) +
  geom_text(aes(label = percent(fa_median_apy_cap_pct, accuracy = 0.1)),
            hjust = -0.2, size = 3.5) +
  scale_fill_manual(values = pos_colors) +
  scale_y_continuous(labels = percent_format(), expand = expansion(mult = c(0, 0.15))) +
  coord_flip() +
  labs(
    title = "Free Agency Replacement Cost by Position",
    subtitle = "Median APY as % of salary cap for FA signings",
    x = NULL, y = "Median APY (% of Cap)"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none")
```

## The Sharpe Ratio

The money shot. This heatmap shows the quasi-Sharpe ratio for every position-tier combination.
Higher values mean better risk-adjusted returns relative to free agency replacement cost.

```{r sharpe-heatmap, fig.height=8}
ggplot(sharpe, aes(x = tier, y = reorder(pos_group, sharpe_linear), fill = sharpe_linear)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(aes(label = round(sharpe_linear, 2)),
            size = 3.5, color = "white", fontface = "bold") +
  scale_fill_gradient2(
    low = "#c62828", mid = "#f9a825", high = "#2e7d32",
    midpoint = 0
  ) +
  labs(
    title = "Quasi-Sharpe Ratio by Position Group and Draft Tier",
    subtitle = "Higher = better risk-adjusted return over FA replacement",
    x = NULL, y = NULL, fill = "Sharpe\nRatio"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 30, hjust = 1),
    panel.grid = element_blank()
  )
```

```{r sharpe-table}
sharpe %>%
  select(pos_group, tier, n, hit_rate, mean_second_contract, fa_replacement, sharpe_linear, sharpe_elite) %>%
  mutate(
    hit_rate = percent(hit_rate, accuracy = 0.1),
    mean_second_contract = percent(mean_second_contract, accuracy = 0.01),
    fa_replacement = percent(fa_replacement, accuracy = 0.01),
    sharpe_linear = round(sharpe_linear, 3),
    sharpe_elite = round(sharpe_elite, 3)
  ) %>%
  gt() %>%
  tab_header(
    title = "Full Sharpe Ratio Breakdown",
    subtitle = "By position group and draft tier"
  ) %>%
  cols_label(
    pos_group = "Position",
    tier = "Tier",
    n = "N",
    hit_rate = "Hit Rate",
    mean_second_contract = "Mean 2nd Contract",
    fa_replacement = "FA Replacement",
    sharpe_linear = "Sharpe (Linear)",
    sharpe_elite = "Sharpe (Elite)"
  )
```

## Sharpe Ratio Curves

How does the Sharpe ratio decay as you move to later draft tiers? Some positions hold value
deep into Day 2 and 3, while others collapse quickly.

```{r sharpe-curves, fig.height=7}
sharpe_key <- sharpe %>%
  filter(pos_group %in% key_positions)

ggplot(sharpe_key, aes(x = tier, y = sharpe_linear, color = pos_group, group = pos_group)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40", linewidth = 0.6) +
  scale_color_manual(values = pos_colors) +
  labs(
    title = "Sharpe Ratio by Draft Tier",
    subtitle = "Key positions | Dashed line = break-even vs. free agency",
    x = NULL, y = "Quasi-Sharpe Ratio (Linear)", color = "Position"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Linear vs Elite-Weighted

The elite-weighted Sharpe ratio penalizes positions where high draft capital rarely produces
truly elite contracts. Comparing the two variants reveals where the "ceiling" matters most.

```{r linear-vs-elite, fig.height=8}
sharpe_long <- sharpe %>%
  filter(pos_group %in% key_positions) %>%
  select(pos_group, tier, sharpe_linear, sharpe_elite) %>%
  pivot_longer(cols = c(sharpe_linear, sharpe_elite),
               names_to = "variant", values_to = "sharpe") %>%
  mutate(variant = recode(variant,
                          "sharpe_linear" = "Linear",
                          "sharpe_elite" = "Elite-Weighted"))

ggplot(sharpe_long, aes(x = tier, y = sharpe, color = pos_group, group = pos_group)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40", linewidth = 0.6) +
  facet_wrap(~variant, ncol = 2) +
  scale_color_manual(values = pos_colors) +
  labs(
    title = "Linear vs Elite-Weighted Sharpe Ratio",
    subtitle = "Key positions | Elite variant rewards positions with high-ceiling outcomes",
    x = NULL, y = "Quasi-Sharpe Ratio", color = "Position"
  ) +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Player Lookup

Recent draft picks with their outcomes and Sharpe context. This table shows the last 5 draft
classes, filtered to players with second-contract data or recent picks still on rookie deals.

```{r player-lookup-table}
max_season <- max(player_lookup$season)
recent <- player_lookup %>%
  filter(season >= max_season - 4) %>%
  arrange(season, pick) %>%
  head(50)

recent %>%
  mutate(
    avg_snap_pct = percent(avg_snap_pct, accuracy = 0.1),
    second_apy_cap_pct = ifelse(second_apy_cap_pct == 0, "--", percent(second_apy_cap_pct, accuracy = 0.01)),
    sharpe_linear = round(sharpe_linear, 2),
    sharpe_elite = round(sharpe_elite, 2)
  ) %>%
  gt() %>%
  tab_header(
    title = "Player Lookup: Recent Draft Classes",
    subtitle = paste0("Seasons ", max_season - 4, "-", max_season, " | Top 50 by pick")
  ) %>%
  cols_label(
    pfr_player_name = "Player",
    pos_group = "Pos",
    season = "Year",
    pick = "Pick",
    tier = "Tier",
    is_hit = "Hit?",
    avg_snap_pct = "Avg Snap %",
    second_apy_cap_pct = "2nd Contract (Cap %)",
    sharpe_linear = "Sharpe (L)",
    sharpe_elite = "Sharpe (E)"
  )
```

## Key Takeaways

```{r takeaways, results='asis'}
# Best top-10 value
best_top10 <- sharpe %>%
  filter(tier == "Top 10") %>%
  arrange(desc(sharpe_linear)) %>%
  slice(1)

# Worst top-10 value
worst_top10 <- sharpe %>%
  filter(tier == "Top 10") %>%
  arrange(sharpe_linear) %>%
  slice(1)

# Biggest elite-weight gainer (position where elite >> linear in top 10)
elite_gain <- sharpe %>%
  filter(tier == "Top 10") %>%
  mutate(elite_delta = sharpe_elite - sharpe_linear) %>%
  arrange(desc(elite_delta)) %>%
  slice(1)

# Best Day 2 value
best_day2 <- sharpe %>%
  filter(tier == "Day 2") %>%
  arrange(desc(sharpe_linear)) %>%
  slice(1)

# Position that stays positive longest
stays_positive <- sharpe %>%
  filter(sharpe_linear > 0) %>%
  group_by(pos_group) %>%
  summarize(deepest_positive = max(as.numeric(tier)), .groups = "drop") %>%
  arrange(desc(deepest_positive)) %>%
  slice(1)

cat(paste0(
  "- **Best top-10 Sharpe ratio**: ", best_top10$pos_group,
  " (", round(best_top10$sharpe_linear, 2), ") -- ",
  "drafting a ", best_top10$pos_group, " in the top 10 offers the best risk-adjusted return.\n\n",

  "- **Worst top-10 Sharpe ratio**: ", worst_top10$pos_group,
  " (", round(worst_top10$sharpe_linear, 2), ") -- ",
  "top-10 ", worst_top10$pos_group, " picks have the weakest return relative to variance.\n\n",

  "- **Biggest elite-weight gainer**: ", elite_gain$pos_group,
  " -- the elite-weighted variant shifts the Sharpe ratio by ",
  round(elite_gain$elite_delta, 2), " vs. the linear version in the top 10, ",
  "because elite ", elite_gain$pos_group, " contracts are disproportionately large.\n\n",

  "- **Best Day 2 value**: ", best_day2$pos_group,
  " (", round(best_day2$sharpe_linear, 2), ") -- ",
  "second-round ", best_day2$pos_group, " picks offer the strongest mid-round return.\n\n",

  "- **Deepest value**: ", stays_positive$pos_group,
  " maintains a positive Sharpe ratio deeper into the draft than any other position, ",
  "suggesting sustained value even with later picks.\n"
))
```
